Enhanced proximity effect in zigzag-shaped Majorana Josephson junctions.

### 2019-02-25 paper/snakemajoranas.tex
wave function is two words

\begin{figure}[!htb]
\centering
\includegraphics[width=0.75\columnwidth]{figures/zigzag.pdf}
\caption{Setup of the straight (top) and the zigzag (bottom) SNS junction.
The yellow areas are the superconductors at a phase difference $\pm \phi / 2$.
The middle area is the semiconductor of width $W$ and has a magnetic field $B$ in the $x$-direction.
The modulation of the zigzag is parameterized by its amplitude $z_y$ and period $z_x$.
Inside the semiconductor we indicate the longest possible trajectory by the red curve.
\label{fig:setup}}
\end{figure}

@@

\begin{figure}[!htb]
\includegraphics[width=\columnwidth]{figures/wavefunctions}
\caption{Density of wavefunctions $\left|\psi\right|^2$ for sizes and geometries.
With (a) a straight system, (b) a saw-toothed zigzag system, (c) a system where the normal region is defined by lines parallel to a sinusoid, and (d) similar to (c) but with disordered edges.
Inside the figure, we indicate the Majorana length (or coherence length) $\xi_\textrm{M}$, the Majorana energy $E_\textrm{M}$ (the energy of the first excited state), and the topological energy gap $E_\textrm{gap}$.
We observe that $\xi_\textrm{M}$ for (a) is almost two orders of magnitude longer and $E_\textrm{gap}$ more than an order of magnitude smaller than for (b), (c), and (d).
The robustness of $E_\textrm{gap}$ and $\xi_\textrm{M}$ across (b), (c), and (d) means that the details of the geometry do not matter for the improvements to occur.
The length of the system is $3.5 z_x=\SI{4550}{\nm}$, the remaining parameters values are written at the end of Sec.~\ref{sec:setup}.\label{fig:wavefunctions}}
\end{figure}

\comment{We calculate the wave functions and find the Majorana lengths by fitting an exponential.}
We model a finite system to compute the Majorana wave function density for different geometries; ribbon, saw-toothed zigzag, parallel curves, and variants of the latter with disordered edges.
By diagonalizing the Hamiltonian, we find the Majorana energy $E_\textrm{M}$ and the energy of the first excited state $E_1$, the lowest and second lowest energies, respectively.
Using the eigenstates, we calculate the Majorana decay length $\xi_\textrm{M}$ by fitting an exponential to the density of a single Majorana wavefunction projected on the $x$-axis.
To determine $\xi_\textrm{M}$, we use a trick and artificially break the particle-hole symmetry on one edge by adding a potential $V \sigma_0 \tau_0$ such that one of the Majorana states is pushed into the continuum and the fit to $\exp{\frac{x}{\xi_\textrm{M}}}$ becomes easier.
In figure~\ref{fig:wavefunctions} the wavefunctions are displayed, superimposed upon their respective geometry.
All systems have identical Hamiltonian parameters in order to compare the effects of the changing geometry meaningfully.

\comment{In a straight system, the Majoranas are very poorly localized.}
For the ribbon system [Fig.~\ref{fig:wavefunctions}(a)], we see that the decay of the density is long compared to the system size.
The wavefunction extends to the center of the system, resulting in overlapping Majoranas and a non-zero energy of the state.
Further, the energy of the Majorana state $E_\textrm{M}$ in the straight system is badly separated from the next lowest lying eigenstate $E_1$.
Altogether, the increase in the topological gap, but also the before mentioned reduction of the group velocity, result in poorly localized Majoranas and minimal topological protection against perturbations.  % TODO: probably put the eq for coherence length here?

\comment{In a zigzag geometry Majoranas are localized within one segment of zigzag.}
We observe that for zigzag systems the Majorana properties improve.
All of the zigzag type geometries display a reduced coherence length and the delocalized nature of the wavefunctions is improved, as is visible through the density plots.
Quantitatively, the improvement in the localization of the Majoranas is also distinctly apparent: the energies of the Majorana states are three orders of magnitude lower than the energy of the second lowest-lying wavefunctions.
As mentioned in section \ref{sec:band_structures}, this can be attributed to the way the energy gap and velocity factor into the Majorana coherence length which scales with these parameters $\xi_\textrm{M}=\hbar\frac{v_\textrm{F}}{\Delta_\textrm{ind}}$.
Additionally, the reduction of the Majorana decay length allows for shorter systems sizes, which is favorable because in an experimental device the coherence length of a material sets a limit for the device size.

\comment{The specific geometry is not important.}
The shape of the wavefunctions does not change significantly depending on the details of the geometry.
The sharp corners of the sawtooth versus the smooth shape of the snakelike system do not have a significant impact on the shape of the wavefunction, and neither do the disordered edges affect the relevant parameters.

@@

\comment{We calculate the topological phase diagram using the gap size.}
In Fig.~\ref{fig:phasediagrams} we plot the energy gap as a function of magnetic field, chemical potential, $E_\textrm{gap}(B_x, \mu)$; and the superconducting phase difference $E_\textrm{gap}(B_x, \phi)$ for both a straight system [(c) and (e)] and a zigzag system [(d) and (f)].
Additionally, we plot the first $\SI{1300}{\nm}$ (one zigzag period) of the wavefunctions [(a) and (b)] at the most optimal point in parameter space for the straight system.
For the straight system we calculate $E_\textrm{gap}$ by performing a binary search in $E$ for the energy at which the propagating modes start to appear.~\cite{nijholt2015orbital}
Additionally, in Fig.~\ref{fig:phasediagrams}(c) we plot the phase boundaries that we find by solving a generalized eigenvalue problem as described in Ref.~\onlinecite{nijholt2015orbital}.
Due to the large size of the zigzag supercell, we are unable to use these full spectrum methods to compute these quantities.
Instead, we calculate $E_\textrm{gap}$ by finding the absolute minimum of the spectrum $E_\textrm{gap}=\min{|E(k)|}$.
By both observing the gap closings and knowing the topological phase for the non-zigzag system, we can infer the topology of the system with zigzag modulation.

>>>>

\begin{figure}[!htb]
\centering
\includegraphics[width=0.75\columnwidth]{figures/zigzag.pdf}
\caption{Setup of the straight (top) and the zigzag (bottom) SNS junction.
The yellow areas are the superconductors at an opposite superconducting phase $\pm \phi / 2$.
The middle area is the semiconductor of width $W$ and has a magnetic field $B$ in the $x$-direction.
The modulation of the zigzag is parameterized by its amplitude $z_y$ and period $z_x$.
Inside the semiconductor we indicate the longest possible trajectory by the red curve.
\label{fig:setup}}
\end{figure}

@@

\begin{figure}[!htb]
\includegraphics[width=\columnwidth]{figures/wave functions}
\caption{Density of wave functions $\left|\psi\right|^2$ for sizes and geometries.
With (a) a straight system, (b) a saw-toothed zigzag system, (c) a system where the normal region is defined by lines parallel to a sinusoid, and (d) similar to (c) but with disordered edges.
Inside the figure, we indicate the Majorana length (or coherence length) $\xi_\textrm{M}$, the Majorana energy $E_\textrm{M}$ (the energy of the first excited state), and the topological energy gap $E_\textrm{gap}$.
We observe that $\xi_\textrm{M}$ for (a) is almost two orders of magnitude longer and $E_\textrm{gap}$ more than an order of magnitude smaller than for (b), (c), and (d).
The robustness of $E_\textrm{gap}$ and $\xi_\textrm{M}$ across (b), (c), and (d) means that the details of the geometry do not matter for the improvements to occur.
The length of the system is $3.5 z_x=\SI{4550}{\nm}$, the remaining parameters values are written at the end of Sec.~\ref{sec:setup}.\label{fig:wave_functions}}
\end{figure}

\comment{We calculate the wave functions and find the Majorana lengths by fitting an exponential.}
We model a finite system to compute the Majorana wave function density for different geometries; ribbon, saw-toothed zigzag, parallel curves, and a variant of the latter with disordered edges.
In figure~\ref{fig:wave_functions} the wave functions are displayed, superimposed upon their respective geometry.
By diagonalizing the Hamiltonian, we find the Majorana energy $E_\textrm{M}$ and by using the corresponding eigenstate of that lowest energy, we get the wave function.
To determine $\xi_\textrm{M}$, we use a trick and artificially break the particle-hole symmetry on one edge by adding a potential $V \sigma_0 \tau_0$ such that one of the Majorana states is pushed into the continuum, then the fit to $\exp{\frac{x}{\xi_\textrm{M}}}$ becomes better.
Using this system, we then find the Majorana decay length $\xi_\textrm{M}$ by fitting an exponential to the density of the single Majorana wave function projected on the $x$-axis.
All systems have identical Hamiltonian parameters in order to compare the effects of the changing geometry meaningfully.

\comment{In a straight system, the Majoranas are very poorly localized.}
For the ribbon system [Fig.~\ref{fig:wave_functions}(a)], we see that the decay of the density is long compared to the system size.
The wave function extends to the center of the system, resulting in overlapping Majoranas and a non-zero energy of the state.
Further, the energy of the Majorana state $E_\textrm{M}$ in the straight system is high in comparison with the topological gap $E_\textrm{gap}$.
Altogether, the increase in the topological gap, but also the before mentioned reduction of the group velocity, result in poorly localized Majoranas and minimal topological protection against perturbations.  % TODO: probably put the eq for coherence length here?

\comment{In a zigzag geometry Majoranas are localized within one segment of zigzag.}
We observe that for zigzag systems the Majorana properties improve.
All of the zigzag type geometries display a reduced coherence length and the delocalized nature of the wave functions is improved, as is visible through the density plots.
Quantitatively, the improvement in the localization of the Majoranas is also distinctly apparent: the energies of the Majorana states are three orders of magnitude lower than the energy of the second lowest-lying wave functions.
As mentioned in section \ref{sec:band_structures}, this can be attributed to the way the energy gap and velocity factor into the Majorana coherence length which scales with these parameters $\xi_\textrm{M}=\hbar\frac{v_\textrm{F}}{\Delta_\textrm{ind}}$.
Additionally, the reduction of the Majorana decay length allows for shorter systems sizes, which is favorable because in an experimental device the coherence length of a material sets a limit for the device size.

\comment{The specific geometry is not important.}
The shape of the wave functions does not change significantly depending on the details of the geometry.
The sharp corners of the sawtooth versus the smooth shape of the snakelike system do not have a significant impact on the shape of the wave function, and neither do the disordered edges affect the relevant parameters.

@@

\comment{We calculate the topological phase diagram using the gap size.}
In Fig.~\ref{fig:phasediagrams} we plot the energy gap as a function of magnetic field, chemical potential, $E_\textrm{gap}(B_x, \mu)$; and the superconducting phase difference $E_\textrm{gap}(B_x, \phi)$ for both a straight system [(c) and (e)] and a zigzag system [(d) and (f)].
Additionally, we plot the first $\SI{1300}{\nm}$ (one zigzag period) of the wave functions [(a) and (b)] at the most optimal point in parameter space for the straight system.
For the straight system we calculate $E_\textrm{gap}$ by performing a binary search in $E$ for the energy at which the propagating modes start to appear.~\cite{nijholt2015orbital}
Additionally, in Fig.~\ref{fig:phasediagrams}(c) we plot the phase boundaries that we find by solving a generalized eigenvalue problem as described in Ref.~\onlinecite{nijholt2015orbital}.
Due to the large size of the zigzag supercell, we are unable to use these full spectrum methods to compute these quantities.
Instead, we calculate $E_\textrm{gap}$ by finding the absolute minimum of the spectrum $E_\textrm{gap}=\min{|E(k)|}$.
By both observing the gap closings and knowing the topological phase for the non-zigzag system, we can infer the topology of the system with zigzag modulation.
